---
title: "Group Assignment"
author: "Zhicong Hu"
date: "12/03/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r loadlibraries, include=FALSE, message=FALSE, warning=FALSE}
library(tidyquant)
library(tidyverse)
library(dplyr)
library(ggplot2)
library(patchwork)
library(Metrics)

library(tseries) 
library(data.table)
library(vars) # vec2var
library(urca)  # ca.jo, ur.df, finland
library(tsDyn) # VECM
```

# Get data

We extract price data of XOM and CVX, from 2018/01/01 to 2020/12/31.

```{r}
data <- tq_get(c("XOM","CVX"), get = "stock.prices", from = "2018-01-01", to = "2020-12-31") %>% 
  dplyr::select(symbol, date, close) %>% 
  mutate(close = round(close, 2))
```

# Visualization

```{r}
ggplot(data, aes(x = date, y = close, color = symbol)) +
  geom_line() +
  theme_minimal() +
  theme(panel.grid.minor.x =  element_blank(), panel.grid.major.x =  element_blank(),
        panel.grid.minor.y =  element_blank()) +
  labs(x = "")
```

# Data Formatting

```{r}
data <- data %>% 
  pivot_wider(names_from = symbol, values_from = close)
```

# Modelling

```{r}
data <- data %>% 
  mutate(spread = XOM-CVX)
```

## Test for stationary

We tested for stationary with the Dickey-Fuller Test for both Exxon Mobil and Chevron's stock price and both are not stationary. Therefore, we can say that stock prices of them are random walks.

```{r}
adf.test(data$CVX)

adf.test(data$XOM)
```

To test for the possibility of cointegration, we also test for stationary for the spread of the two stock prices (difference of the two stock prices) and we can see that their spread is actually stationary, meaning that their price is mean-reverting. The z-score of our test is as high as -4.161.

```{r warning=FALSE}
adf.test(data$CVX-data$XOM)

ggplot(data, aes(x = date, y = spread)) +
  geom_line() +
  theme_minimal() +
  theme(legend.position = "none")
```

## VAR

###Individual

We use AIC to select the lag we want to include for prediction.

```{r}
df <- data %>% dplyr::select(CVX, XOM)

VARselect(df, lag.max = 10, type = "const")
```

Therefore, we select lag length p = 8 for VAR since it has the lowest AIC.

Now, we run the simple VAR model without error correction term.

```{r}
var.model <- VAR(df, p = 8, type = "const")
summary(var.model)
```

With our model built, we compare our predicted prices with actual prices of between year 2021.

```{r fig.height=8, fig.width=8}
future_data <- tq_get(c("CVX","XOM"), get = "stock.prices", from = "2020-12-21", to = "2021-12-31") %>% 
  dplyr::select(symbol, date, close) %>% 
  mutate(close = round(close, 2)) %>% 
  pivot_wider(names_from = symbol, values_from = close)

var.pred <- data.frame(CVX_pre = NA, CVX_lower = NA, CVX_higher = NA, 
                       XOM_pre = NA, XOM_lower = NA, XOM_higher = NA)
for (i in 1:(nrow(future_data)-8)) {
  temp_model <- VAR(rbind(df, future_data[8:(7+i),c("CVX","XOM")]), p = 8, type = "const")
  var.pred[i,"CVX_pre"] <- predict(temp_model, exoPred = NULL, n.ahead=1)$fcst$CVX[1]
  var.pred[i,"CVX_lower"] <- predict(temp_model, exoPred = NULL, n.ahead=1)$fcst$CVX[2]
  var.pred[i,"CVX_higher"] <- predict(temp_model, exoPred = NULL, n.ahead=1)$fcst$CVX[3]
  var.pred[i,"XOM_pre"] <- predict(temp_model, exoPred = NULL, n.ahead=1)$fcst$XOM[1]
  var.pred[i,"XOM_lower"] <- predict(temp_model, exoPred = NULL, n.ahead=1)$fcst$XOM[2]
  var.pred[i,"XOM_higher"] <- predict(temp_model, exoPred = NULL, n.ahead=1)$fcst$XOM[3]
}

(future_data[9:nrow(future_data),] %>% 
  cbind(XOM_pre = var.pred$XOM_pre, XOM_lower = var.pred$XOM_lower, XOM_higher = var.pred$XOM_higher) %>% 
  ggplot(aes(x = date)) +
    geom_line(aes(y = XOM, color = "Actual")) +
    geom_line(aes(y = XOM_pre, color = "Predicted")) +
    geom_line(aes(y = XOM_lower), color = "darkred", linetype = "dotted", alpha = 0.7) +
    geom_line(aes(y = XOM_higher), color = "darkred", linetype = "dotted", alpha = 0.7) +
    scale_color_manual("",
                       breaks = c("Actual","Predicted"),
                       values = c("blue","coral")) +
    theme_minimal() +
    labs(title = "Predicted Price of XOM with VAR vs Actual in 2021", x = "", y = "")) / 
  (future_data[9:nrow(future_data),] %>% 
  cbind(CVX_pre = var.pred$CVX_pre, CVX_lower = var.pred$CVX_lower, CVX_higher = var.pred$CVX_higher) %>% 
  ggplot(aes(x = date)) +
    geom_line(aes(y = CVX, color = "Actual")) +
    geom_line(aes(y = CVX_pre, color = "Predicted")) +
    geom_line(aes(y = CVX_lower), color = "darkred", linetype = "dotted", alpha = 0.7) +
    geom_line(aes(y = CVX_higher), color = "darkred", linetype = "dotted", alpha = 0.7) +
    scale_color_manual("",
                       breaks = c("Actual","Predicted"),
                       values = c("blue","coral")) +
    theme_minimal() +
    labs(title = "Predicted Price of CVX with VAR vs Actual in 2021", x = "", y = ""))
  
```

As we can see, the actual prices falls out of our predictions' confidence intervals. This actually aligns with our previous findings that the stock prices are random walks and hence is not predictable. However, as we mentioned, the spread of the two prices is actually stationary and we will move on to building the model that is able to capture the this information.

### Spread

Now, we build VAR model for prediction of the spread.

```{r eval=FALSE}
df.diff <- diff(as.matrix(df), lag = 1)
# estimation
vare_diff <- lineVar(df.diff, lag = 8, include = "const", model = "VAR")
vare_diff

varf_diff <- data.frame(CVX_pre = NA, XOM_pre = NA)
for (i in 1:(nrow(future_data)-8)) {
  varf_diff[i,"CVX_pre"] <- predict(vare_diff, newdata = future_data[i:(i+7),c("CVX","XOM")], exoPred = NULL,
                                   n.ahead=1)[1]
  varf_diff[i,"XOM_pre"] <- predict(vare_diff, newdata = future_data[i:(i+7),c("CVX","XOM")], exoPred = NULL,
                                   n.ahead=1)[2]
}
varf_diff$CVX_pre <- cumsum(varf_diff$CVX_pre) + future_data$CVX[8]
varf_diff$XOM_pre <- cumsum(varf_diff$XOM_pre) + future_data$XOM[8]


(tq_get(c("XOM"), get = "stock.prices", from = "2010-01-01", to = "2010-12-31") %>% 
  dplyr::select(symbol, date, close) %>% 
  mutate(close = round(close, 2)) %>% 
  pivot_wider(names_from = symbol, values_from = close) %>% 
  cbind(XOM_pre = df$XOM[nrow(df)] + cumsum(varf_diff$fcst$XOM[,1]), 
        XOM_lower = df$XOM[nrow(df)] + cumsum(varf_diff$fcst$XOM[,1]) + varf_diff$fcst$XOM[,2], 
        XOM_higher = df$XOM[nrow(df)] + cumsum(varf_diff$fcst$XOM[,1]) + varf_diff$fcst$XOM[,3]) %>% 
  ggplot(aes(x = date)) +
    geom_line(aes(y = XOM, color = "Actual")) +
    geom_line(aes(y = XOM_pre, color = "Predicted")) +
    geom_line(aes(y = XOM_lower), color = "darkred", alpha = 0.7, linetype = "dotted") +
    geom_line(aes(y = XOM_higher), color = "darkred", alpha = 0.7, linetype = "dotted") +
    scale_color_manual("",
                       breaks = c("Actual","Predicted"),
                       values = c("darkgreen","darkblue")) +
    theme_minimal() +
    labs(title = "Predicted Price of XOM with VAR vs Actual in 2010", x = "", y = "")) / 
  (tq_get(c("CVX"), get = "stock.prices", from = "2010-01-01", to = "2010-12-31") %>% 
  dplyr::select(symbol, date, close) %>% 
  mutate(close = round(close, 2)) %>% 
  pivot_wider(names_from = symbol, values_from = close) %>% 
  cbind(CVX_pre = df$CVX[nrow(df)] + cumsum(varf_diff$fcst$CVX[,1]), 
        CVX_lower = df$CVX[nrow(df)] + cumsum(varf_diff$fcst$CVX[,1]) + varf_diff$fcst$CVX[,2], 
        CVX_higher = df$CVX[nrow(df)] + cumsum(varf_diff$fcst$CVX[,1]) + varf_diff$fcst$CVX[,3]) %>% 
  ggplot(aes(x = date)) +
    geom_line(aes(y = CVX, color = "Actual")) +
    geom_line(aes(y = CVX_pre, color = "Predicted")) +
    geom_line(aes(y = CVX_lower), color = "darkred", alpha = 0.7, linetype = "dotted") +
    geom_line(aes(y = CVX_higher), color = "darkred", alpha = 0.7, linetype = "dotted") +
    scale_color_manual("",
                       breaks = c("Actual","Predicted"),
                       values = c("darkgreen","darkblue")) +
    theme_minimal() +
    labs(title = "Predicted Price of CVX with VAR vs Actual in 2010", x = "", y = ""))


```

## VECM

Showing that our VAR model does not work well with two random walks and that the spread of the two random walks are stationary, we can include the spread into our model to get a VECM model.

```{r}
#========================================================
# Cointegration Test
#========================================================

#———————————————-
# Johansen Cointegration Procedure
#———————————————-
# ecdet  = ‘none’  for no intercept 
#          ‘const’ for constant term
#          ‘trend’ for trend variable 
#          in cointegration
# type   =  eigen or trace test
# K      =  lag order of VAR
# spec   = "transitory" or "longrun"
# season = centered seasonal dummy (4:quarterly)
# dumvar = another dummy variables
#———————————————-

coint_ca.jo <- ca.jo(
    df, ecdet = "const", type  = "eigen", K = 8, 
    spec = "transitory", season = NULL, dumvar = NULL)
summary(coint_ca.jo)
```

We can interpret our results of test-statistics and critical values. The test of r = 0, which has the null hypothesis that there is no presence of cointegration, we can reject with a level of significance of 5% but we are unable to reject the test of r <= 1. Therefore, we conclude that cointegration exist with a r = 1 or the rank of the matrix is 1, i.e. there is one cointegrating vector.

Since we have prove that there are cointegration, we can move on to building the VECM model.

```{r warning=FALSE}
#————————————————
# VECM estimation
#————————————————
# VECM(data, lag, r = 1, 
#      include = c("const", "trend", "none", "both"),
#      beta = NULL, estim = c("2OLS", "ML"), 
#      LRinclude = c("none", "const","trend", "both"), 
#      exogen = NULL)
#————————————————
  
VECM_tsDyn <- VECM(df, lag = 8, r = 1, estim = "2OLS",
                   LRinclude = "const", exogen = NULL)
summary(VECM_tsDyn)
```

As we can see from our model's output, the Error Correction Term is significant for CVX but not for XOM. This suggest that CVX is the follower while XOM is the leader, XOM's stock prices leads and when the spread between the prices becomes larger, CVX will have to follow the price of XOM and adjust their prices accordingly.


Forecast

```{r}
#========================================================
# Forecast
#========================================================

#———————————————-
# Forecast from VECM() in tsDyn
#———————————————-

VECM_pred_tsDyn <- data.frame(CVX = NA, XOM = NA)
for (i in 1:(nrow(future_data)-8)) {
  VECM_pred_tsDyn[i,"CVX"] <- predict(VECM_tsDyn, newdata = future_data[i:(i+8),c("CVX","XOM")], exoPred = NULL, n.ahead=1)[1]
  VECM_pred_tsDyn[i,"XOM"] <- predict(VECM_tsDyn, newdata = future_data[i:(i+8),c("CVX","XOM")], exoPred = NULL, n.ahead=1)[2]
}
```

```{r fig.height=8, fig.width=8}
(future_data[9:nrow(future_data),] %>% 
  cbind(XOM_pre = VECM_pred_tsDyn[,"XOM"], XOM_pre_VAR = var.pred$XOM_pre) %>% 
  ggplot(aes(x = date)) +
    geom_line(aes(y = XOM, color = "Actual")) +
    geom_line(aes(y = XOM_pre, color = "VECM Predicted")) +
    geom_line(aes(y = XOM_pre_VAR, color = "VAR Predicted")) +
    scale_color_manual("",
                       breaks = c("Actual","VECM Predicted", "VAR Predicted"),
                       values = c("darkgreen","darkblue", "darkred")) +
    theme_minimal() +
    labs(title = "Predicted Price of XOM with VECM vs VAR vs Actual in 2021", x = "", y = "")) / 
  (future_data[9:nrow(future_data),] %>% 
  cbind(CVX_pre = VECM_pred_tsDyn[,"CVX"], CVX_pre_VAR = var.pred$CVX_pre) %>% 
  ggplot(aes(x = date)) +
    geom_line(aes(y = CVX, color = "Actual")) +
    geom_line(aes(y = CVX_pre, color = "VECM Predicted")) +
    geom_line(aes(y = CVX_pre_VAR, color = "VAR Predicted")) +
    scale_color_manual("",
                       breaks = c("Actual","VECM Predicted", "VAR Predicted"),
                       values = c("darkgreen","darkblue", "darkred")) +
    theme_minimal() +
    labs(title = "Predicted Price of CVX with VECM vs VAR vs Actual in 2021", x = "", y = ""))
```

We evaluate the performance of predictive model by out-of-sample RMSE.

```{r}
data.frame(
  RMSE = c("CVX","XOM"),
  VAR = c(rmse(future_data[9:nrow(future_data),]$XOM, var.pred$XOM_pre), 
          rmse(future_data[9:nrow(future_data),]$CVX, var.pred$CVX_pre)),
  VECM = c(rmse(future_data[9:nrow(future_data),]$XOM, VECM_pred_tsDyn[,"XOM"]),
           rmse(future_data[9:nrow(future_data),]$CVX, VECM_pred_tsDyn[,"CVX"]))
)
```

We evaluate the performance of predictive model by out-of-sample R-squared (the deviation that our model is able to explain).

```{r}
data.frame(
  R_squared = c("CVX","XOM"),
  VAR = c(cor(future_data[9:nrow(future_data),]$XOM, var.pred$XOM_pre)^2, 
          cor(future_data[9:nrow(future_data),]$CVX, var.pred$CVX_pre)^2),
  VECM = c(cor(future_data[9:nrow(future_data),]$XOM, VECM_pred_tsDyn[,"XOM"])^2,
           cor(future_data[9:nrow(future_data),]$CVX, VECM_pred_tsDyn[,"CVX"])^2)
)
```

The out-of-sample R-squared of our VECM model is extremely high as well.

# Time Period

We decide to look at our data from different time periods. For comparison purposes, all of our models will be trained with 3 years of data and tested with 1 year of data following the 3 year period.

First, we trained our model with 2010-2012 data to predict the prices of 2013.

```{r warning=FALSE}
VECM_2010 <- VECM(tq_get(c("CVX","XOM"), get = "stock.prices", from = "2010-01-01", to = "2012-12-31") %>% dplyr::select(symbol, date, close) %>% 
                    mutate(close = round(close, 2)) %>% pivot_wider(names_from = symbol, values_from = close) %>% dplyr::select(CVX, XOM), 
                  lag = 8, r = 1, estim = "2OLS", LRinclude = "const", exogen = NULL)
summary(VECM_2010)

future_data_2013 <- tq_get(c("CVX","XOM"), get = "stock.prices", from = "2012-12-19", to = "2013-12-31") %>% 
  dplyr::select(symbol, date, close) %>% 
  mutate(close = round(close, 2)) %>% 
  pivot_wider(names_from = symbol, values_from = close)

VECM_pred_2013 <- data.frame(CVX = NA, XOM = NA)
for (i in 1:(nrow(future_data_2013)-8)) {
  VECM_pred_2013[i,"CVX"] <- predict(VECM_2010, newdata = future_data_2013[i:(i+8),c("CVX","XOM")], exoPred = NULL, n.ahead=1)[1]
  VECM_pred_2013[i,"XOM"] <- predict(VECM_2010, newdata = future_data_2013[i:(i+8),c("CVX","XOM")], exoPred = NULL, n.ahead=1)[2]
}
```

Then, we trained our model with 2014-2016 data to predict the prices of 2017.


```{r warning=FALSE}
VECM_2014 <- VECM(tq_get(c("XOM", "CVX"), get = "stock.prices", from = "2014-01-01", to = "2016-12-31") %>% dplyr::select(symbol, date, close) %>% 
                    mutate(close = round(close, 2)) %>% pivot_wider(names_from = symbol, values_from = close) %>% dplyr::select(CVX, XOM), 
                  lag = 8, r = 1, estim = "2OLS", LRinclude = "const", exogen = NULL)
summary(VECM_2014)

future_data_2017 <- tq_get(c("CVX","XOM"), get = "stock.prices", from = "2016-12-20", to = "2017-12-31") %>% 
  dplyr::select(symbol, date, close) %>% 
  mutate(close = round(close, 2)) %>% 
  pivot_wider(names_from = symbol, values_from = close)

VECM_pred_2017 <- data.frame(CVX = NA, XOM = NA)
for (i in 1:(nrow(future_data_2017)-8)) {
  VECM_pred_2017[i,"CVX"] <- predict(VECM_2014, newdata = future_data_2017[i:(i+8),c("CVX","XOM")], exoPred = NULL, n.ahead=1)[1]
  VECM_pred_2017[i,"XOM"] <- predict(VECM_2014, newdata = future_data_2017[i:(i+8),c("CVX","XOM")], exoPred = NULL, n.ahead=1)[2]
}
```


```{r}
data.frame(
  VECM = c("CVX","XOM"),
  Year2013 = c(rmse(future_data_2013[9:nrow(future_data_2013),]$CVX, VECM_pred_2013[,"CVX"]),
               rmse(future_data_2013[9:nrow(future_data_2013),]$XOM, VECM_pred_2013[,"XOM"])),
  Year2017 = c(rmse(future_data_2017[9:nrow(future_data_2017),]$CVX, VECM_pred_2017[,"CVX"]),
               rmse(future_data_2017[9:nrow(future_data_2017),]$XOM, VECM_pred_2017[,"XOM"])),
  Year2021 = c(rmse(future_data[9:nrow(future_data),]$XOM, VECM_pred_tsDyn[,"XOM"]),
               rmse(future_data[9:nrow(future_data),]$CVX, VECM_pred_tsDyn[,"CVX"]))
)
```

We can see that however, our prediction is despite being quite accurate, are prone to time period changes.

# Leader Analysis

As we can see from above analysis, between 2018-2020, XOM was the leader while CVX was the follower. We are interest to see who is the leadership of the energy industry. We extract price data of all component stocks in the S&P500 Energy Sector Index, dating from 2020-01-01 till 2021-12-31, for analysis.


```{r}
energy_stocks <- c("APA","BKR","COP","CTRA","CVX","DVN","EOG","FANG","HAL","HES","KMI","MPC","MRO","OKE","OXY","PSX","PXD","SLB",
                   "VLO","WMB","XOM")

data <- tq_get(energy_stocks, get = "stock.prices", from = "2020-01-01", to = "2021-12-31") %>% 
  dplyr::select(symbol, date, close) %>% 
  mutate(close = round(close, 2)) %>% 
  pivot_wider(names_from = symbol, values_from = close) %>% 
  dplyr::select(-date)

head(data)
```

The one shown are the followers.

```{r warning=FALSE}
leader_analysis <- as.data.frame(matrix(nrow = length(data), ncol = length(data), dimnames = list(colnames(data),colnames(data))))

for (i in 1:length(data)) {
  for (j in 1:length(data)) {
    if (j > i) {
      temp_data <- as.matrix(cbind(data[,i],data[,j]))
      temp_model <- summary(VECM(temp_data, lag = 8, r = 1, estim = "2OLS", LRinclude = "const", exogen = NULL))
      if (is_empty(colnames(temp_model$model)[1:2][grepl("\\*", temp_model$bigcoefficients[,1])])) {
        leader_analysis[i,j] <- NA
      } else {
        leader_analysis[i,j] <- paste(colnames(temp_model$model)[1:2][grepl("\\*", temp_model$bigcoefficients[,1])], collapse=", ")
      }
    }
  }
}
```

